---
title: 'Project 1: What did the presidents say at their inauguation?'
author: "Mengqi Chen"
output:
  html_document:
    df_print: paged
  word_document: default
---
##### The inaugural address has been a tradition set by the first president George Washington. It is made every four years by all the presidents before they take the office. Those political speeches are very important since the president’s fundamental political values and outline of his policy will be pronounced. Inaugurals in US provide a good indication of forthcoming political agenda. 
##### An artical is made of sentences and a sentence consist of words. This report focus on analysing the basic part- words of inaugural speeches including the length of speeches, the lexical diversity and the frequency of words. 


##### Prepare: Install packages; Set plot theme
```{r,echo=FALSE,warning=F}
packages.used=c("tm", "wordcloud","RColorBrewer", "dplyr", "tidytext","ggplot2")
install.packages(packages.used,repos = 'http://cran.us.r-project.org')
packages.needed=setdiff(packages.used, 
                        intersect(installed.packages()[,1], 
                                  packages.used))
# install additional packages
if(length(packages.needed)>0){
  install.packages(packages.needed, dependencies = TRUE)
}
library(tm)
library(wordcloud)
library(RColorBrewer)
library(dplyr)
library(tidytext)
library(ggplot2)

plotTheme <- function(base_size = 12) {
theme(
text = element_text( color = "black"),
plot.title = element_text(size = 10,colour = "black",hjust=0.5),
plot.subtitle = element_text(face="italic"),
plot.caption = element_text(hjust=0),
axis.ticks = element_blank(),
panel.background = element_blank(),
panel.grid.major = element_line("grey80", size = 0.1),
panel.grid.minor = element_blank(),
strip.background = element_rect(fill = "grey80", color = "white"),
strip.text = element_text(size=8),
axis.title = element_text(size=5),
axis.text = element_text(size=5),
axis.title.x = element_text(hjust=1),
axis.title.y = element_text(hjust=1),
plot.background = element_blank(),
legend.background = element_blank(),
legend.title = element_text(colour = "black", face = "bold"),
legend.text = element_text(colour = "black", face = "bold"))
}
```
 
# Part 1: Numbers of words
#### 1.1 The number of words used by each presidents
```{r,echo=FALSE,warning=F}
# Import csv data from the local file which contains name of presidents, speeches' file name, which term, party and how many words used in the speech.
inau.info <- read.csv("/Users/mengqichen/Downloads/Github/Spring2018-Project1-mc4396/data/InaugurationInfo.csv", header = T, as.is = T)
names(inau.info) <- c("name","file","term","party","words")
year <- seq(1789, 2017, 4)
inau.info$year <- as.numeric(year)
inau.info$words <- as.numeric(inau.info$words)
head(inau.info)
library(ggplot2)
ggplot(inau.info,aes(x=year,y=words))+geom_bar(stat = "identity",fill = "#0C8DC4",width=1.3,position=position_dodge(0.7))+ coord_flip()+theme(axis.text.x = element_text(vjust=1,angle=90))+plotTheme()+geom_text(aes(label=name), vjust=0,angle=0,size=2.5,hjust=0)+labs(title="How do the speeches'length change by years?",caption="Longest:William Henry Harrison;Shortest: George Washington( 2rd)")+ylim(c(0,10000))
```

From the graph, President William Henry Harrison had the longest speech over 8000 words and George Wahington had the shortest one only 135 words in his second term.

Highly influential Presdents like Abraham Lincoln, Franklin Delano Roosevelt, George Washington, Renold Reagan seem have shorter speech comparatively.

Some Presidents who did not well, according to the US News, include Herbert Hoover, William Henry Harrison, Ulysses S.Grant, John Tyler. It seems like that most of them had longer speech from the above plot.


```{r,echo=FALSE}
# Import data concerning the content of speech; remove all commas, punctuations and white spaces, lower letters. Then how many different words each president used and words' frequency.

file.path="/Users/mengqichen/Downloads/Github/Spring2018-Project1-mc4396/data/InauguralSpeeches/"
# list all files in this folder
speeches=list.files(path = file.path, pattern = "*.txt") 
head(speeches,3)
length(speeches)
prex.out=substr(speeches, 6, nchar(speeches)-4)
# represneting and computing on corpora
file.all<-Corpus(DirSource(file.path))
```

```{r,echo=F,warning=F}
# next we remove all punctuations and lower all letters. Then we compute the [Document-Term Matrix (DTM)]
file.all<-tm_map(file.all, stripWhitespace)
file.all<-tm_map(file.all, content_transformer(tolower))
file.all<-tm_map(file.all, removeWords, stopwords("english"))
file.all<-tm_map(file.all, removeWords, character(0))
file.all<-tm_map(file.all, removePunctuation)

dtm <- DocumentTermMatrix(file.all,
                          control = list(weighting = function(x)
                                             weightTfIdf(x, 
                                                         normalize =FALSE),
                                         stopwords = TRUE))
file.dtm=tidy(dtm)
```

#### 1.2 LEXICAL DIVERSITY AND NUMBER OF WORDS

##### Lexical Diversity is a measure of the vocabulary of a text. the higher the lexical diversity is, the higher is the vocabulary. 
##### Diversity(rate)= unique(words)/length(words)
##### Assumption: is there any relationship between lexical diversity and number of words? 
```{r, warning=F,echo=F}
word.rebundant <- c()
for (i in 1:58) {
  re <- nrow(file.dtm[file.dtm$document==i,])
  word.rebundant <- c(word.rebundant,re)
}
order <- c(19,20,11,12,56,57,26,35,58,42,43,37,38,39,40,17,51,1,2,54,55,25,27,41,36,18,24,15,6,7,8,9,48,3,44,10,45,13,46,47,49,50,23,30,4,5,21,22,34,14,31,52,53,28,29,32,33,16)
wr.df <- data.frame(prex.out,word.rebundant,order,inau.info$words,inau.info$year)
wr.df <- wr.df[order(order),]
wr.df$diversity <- wr.df$word.rebundant/inau.info$words

ggplot(wr.df,aes(x=year,y=diversity))+geom_line()+geom_point()+theme(axis.text.x = element_text(vjust=1,angle=90))+labs(x="Year",y="Lexical Diversity",title="Lexical Diversity")+plotTheme()+geom_text(aes(label=prex.out),hjust=-0.1, vjust=0,size=2.5,angle=90,check_overlap = T)+ylim(c(0,0.9))
```

Almost all inaugural speeches have lexical diversity, less than 0.5, arond 0.3. The highest one was George Washington's 2rd inaugural speech, which is also the shortest speech. Willian Henry Harrison had the lowest lexical diversity. Is there anyrelationship between the lexical diversity and the number of words?

```{r,echo=FALSE,warning=F}
ggplot(wr.df,aes(x=inau.info.words,y=diversity))+geom_point(alpha=1/3,aes(size=diversity))+geom_smooth(se=F)+plotTheme()+labs(title="Plot Shows the Variation Between Lexical Diversity and Length of Words ")
```

It seems that there is no trend or relationship between lexical diversity and the number of words. The assumption is suffocated.

# Part2: the frequency of words (the overall perspective, the main parties, the interactive comparsion)
#### 2.1 Overal perspective
```{r, fig.height=3, fig.width=3,echo=FALSE,warning=F}
tdm.all<-TermDocumentMatrix(file.all)
tdm.tidy=tidy(tdm.all)
tdm.overall=summarise(group_by(tdm.tidy, term), sum(count))

wordcloud(tdm.overall$term, tdm.overall$`sum(count)`,
          scale=c(5,0.5),
          max.words=100,
          min.freq=1,
          random.order=FALSE,
          rot.per=0.3,
          use.r.layout=T,
          random.color=FALSE,
          colors=brewer.pal(9,"Reds"))
```

The large ones are Will, government, people, can... Will is future tense as we known, the whole speech is talking about what the president is going to do. Goverment and people，actually, the relationship of government and people at different levels is one of the prevailing theme of American history. People at large select goverment and in return the government would take actions to make this country better, to satisfy public demands and solve social crisis.

#### 2.2 From two parties perspective: Repluclican(25 files) and Democratic(22 files)
##### Except George Washington(no party), other parties like whig, Fedralist, most of time in US are dominated by Repuclican or Democratic. Is there any difference in frequent words?
```{r,echo=FALSE,warning=F}
# Load speeches of two parties
##Republican corpus
Republic.txt <- file.path("/Users/mengqichen/Downloads/Github/Spring2018-Project1-mc4396/data/re")
Republic <- Corpus(DirSource(Republic.txt))
Republic

##Democratic corpus
Democratic.txt <- file.path("/Users/mengqichen/Downloads/Github/Spring2018-Project1-mc4396/data/de")
Democratic <- Corpus(DirSource(Democratic.txt))
Democratic
```

```{r,fig.height=4,fig.width=4,echo=FALSE,warning=F}
#Republic wordcloud
Republic<-tm_map(Republic, stripWhitespace)
Republic<-tm_map(Republic, content_transformer(tolower))
Republic<-tm_map(Republic, removeWords, stopwords("english"))
Republic<-tm_map(Republic, removeWords, character(0))
Republic<-tm_map(Republic, removePunctuation)

tdm.Rpbl <- TermDocumentMatrix(Republic)   
tdm.tidyRpbl=tidy(tdm.Rpbl)
tdm.overallRpbl=summarise(group_by(tdm.tidyRpbl, term), sum(count))

wordcloud(tdm.overallRpbl$term, tdm.overallRpbl$`sum(count)`, max.words=60, random.order=FALSE, random.color=FALSE,rot.per=0, colors=brewer.pal(5,"Reds"))

# Democratic wordcrowd
Democratic<-tm_map(Democratic, stripWhitespace)
Democratic<-tm_map(Democratic, content_transformer(tolower))
Democratic<-tm_map(Democratic, removeWords, stopwords("english"))
Democratic<-tm_map(Democratic, removeWords, character(0))
Democratic<-tm_map(Democratic, removePunctuation)

tdm.Dmcr <- TermDocumentMatrix(Democratic)   
tdm.tidyDmcr=tidy(tdm.Dmcr)
tdm.overallDmcr=summarise(group_by(tdm.tidyDmcr, term), sum(count))

wordcloud(tdm.overallDmcr$term, tdm.overallDmcr$`sum(count)`, max.words=60, random.order=FALSE, random.color=FALSE,rot.per=0, colors=brewer.pal(5,"Greens")) 
```

```{r,fig.height=7, fig.width=7,warning=F,echo=F}
# Repuclic
dtm.Rpbl <- DocumentTermMatrix(Republic) 
freq.Rpbl <- colSums(as.matrix(dtm.Rpbl))
length(freq.Rpbl)

#sort the terms based on frequency in a descending order
ord.Rpbl <- order(freq.Rpbl, decreasing = TRUE)
#view top 30
freq.Rpbl[head(ord.Rpbl, n=30L)]

wf.Rpbl <- data.frame(word=names(freq.Rpbl), freq=freq.Rpbl)
library(ggplot2)
p.Rpbl <- ggplot(subset(wf.Rpbl, freq>76), aes(reorder(word, -freq),freq))    
p.Rpbl <- p.Rpbl + geom_bar(stat="identity",fill ="Pink")
p.Rpbl <- p.Rpbl + theme(axis.text.x=element_text(angle=45,size=10, hjust=1)) 
p.Rpbl <- p.Rpbl + xlab('Top 30 words') + ylab('Frequency')
p.Rpbl+plotTheme()

# Democratic
dtm.Dmcr <- DocumentTermMatrix(Democratic) 
freq.Dmcr <- colSums(as.matrix(dtm.Dmcr))
length(freq.Dmcr)

#sort the terms based on frequency in a descending order
ord.Dmcr <- order(freq.Dmcr, decreasing = TRUE)
#view top 30
freq.Dmcr[head(ord.Dmcr, n=30L)]
wf.Dmcr <- data.frame(word=names(freq.Dmcr), freq=freq.Dmcr)
p.Dmcr <- ggplot(subset(wf.Dmcr, freq>55), aes(reorder(word, -freq),freq))    
p.Dmcr <- p.Dmcr + geom_bar(stat="identity",fill = "#56B4E9")
p.Dmcr <- p.Dmcr + theme(axis.text.x=element_text(angle=45,size=10, hjust=1))
p.Dmcr <- p.Dmcr + xlab('Top 30 words') + ylab('Frequency')
p.Dmcr+plotTheme()
```

Those show the similarities and differences among words used by presidents from two parties. There is not so much difference between frequent words used by two parties, still mention “people” and “government” most.Even though two parties, all presidents were commiting themselves to the people, constitution and the nation, try to bring changes and make the country better. 

#### 2.3 Interactive wordcloud( BarackObama, DonaldJTrump)
```{r,echo=FALSE,warning=F}
library(shiny)

shinyApp(
    ui = fluidPage(
      fluidRow(style = "padding-bottom: 20px;",
        column(4, selectInput('speech1', 'Speech 1', speeches, selected=speeches[5])),
        column(4, selectInput('speech2', 'Speech 2', speeches, selected=speeches[9])),
        column(4, sliderInput('nwords', 'Number of words', 3, min = 20, 
                              max = 200, value=100, step = 20))
      ),
      fluidRow(
        plotOutput('wordclouds', height = "400px")
      )
    ),

    server = function(input, output, session) {
      # Combine the selected variables into a new data frame
      selectedData <- reactive({
        list(dtm.term1=file.dtm$term[file.dtm$document==as.character(which(speeches == input$speech1))],
             dtm.count1=file.dtm$count[file.dtm$document==as.character(which(speeches == input$speech1))],
             dtm.term2=file.dtm$term[file.dtm$document==as.character(which(speeches == input$speech2))],
             dtm.count2=file.dtm$count[file.dtm$document==as.character(which(speeches == input$speech2))])
      })

      output$wordclouds <- renderPlot(height = 400, {
        par(mfrow=c(1,2), mar = c(0, 0, 3, 0))
        wordcloud(selectedData()$dtm.term1, 
                  selectedData()$dtm.count1,
              scale=c(4,0.5),
              max.words=input$nwords,
              min.freq=1,
              random.order=FALSE,
              rot.per=0,
              use.r.layout=FALSE,
              random.color=FALSE,
              colors=brewer.pal(10,"Reds"), 
            main=input$speech1)
        wordcloud(selectedData()$dtm.term2, 
                  selectedData()$dtm.count2,
              scale=c(4,0.5),
              max.words=input$nwords,
              min.freq=1,
              random.order=FALSE,
              rot.per=0,
              use.r.layout=FALSE,
              random.color=FALSE,
              colors=brewer.pal(10,"Blues"), 
            main=input$speech2)
      })
    },

    options = list(height = 600)
)
```
This part is an interactive picture that we have seen in last class. Let's see wordclouds of Obama(Red one) 1st inaugural speech and Trump(Green one). He mentioned crisis, job, workers a lot, which reminds us the background when he came to power. It was 2009, the finacial crisis, economic downturn, high unemployment rate. Generation, Obama's polling strategy, one of them, was focused on new generation and he highly valued young people's involvement in community and politics. So did the "women". "Iraq", Bush pronounced to figh terrorisim since 911 and started the Iraq war, which caused a huge financial burden, sacrificed lots of money and soldiers. It turnde out the war ended in 2011. 
Trump mainly talked about "American Dream". The interesting thing one is Obama was the frequent word- he expressed gratefulness to Obama. Back, he insists to bring factories back to domestic, bring back jobs, bring back dreams. 










